### 2024-03-27 sections/5_validation.tex

\section{Validation with $\xi_{\pm}$}
\label{sec:validation}

\input{tables/ia_models}

%------------------------------
\subsection{NLA model}
\label{subsubsec:sigma_G}

See Fig. \ref{fig:xi_NLA}

\input{figures/shear_corr_funcs/xi_nla}

%------------------------------
\subsection{Extended-NLA model}
\label{subsec:deltaNLA}

See Fig. \ref{fig:xi_deltaNLA}.

\input{figures/shear_corr_funcs/xi_delta_nla}

\input{figures/shear_corr_funcs/xi_delta_nla_rescaled}

We verified that it works equally well for $b_{\rm TA}$ = 1.0 and 2.0.

%------------------------------

Using this, we $b_{\rm TA}$-rescale our $b_{\rm TA}=1.0$ mock to $b_{\rm TA}=2.0$ and to $0.0$ mocks. Results are shown in Fig. \ref{fig:xi_deltaNLA_rescaled} for one of the tomographic bin. 
It works reasonably well, but lower redshift are less accurate due to the increased  shot noise (not shown), hence we do not recommend using this.

%------------------------------
\subsection{TT model}
\label{subsec:TT}

See Fig. \ref{fig:xi_TT}

\input{figures/shear_corr_funcs/xi_tt}

\input{figures/shear_corr_funcs/xi_delta_tt}

To match the theory, we need to apply the following calibration: $\epsilon^{\rm IA, TT}(z<0.5) \rightarrow \epsilon^{\rm IA, TT} /2.5$.

%------------------------------
\subsection{Extended-TT model}
\label{subsec:TT}

See Fig. \ref{fig:xi_deltaTT}

To match the TT theory, we need to apply the following calibration: $\epsilon^{\rm IA, TT}(z<0.5) \rightarrow \epsilon^{\rm IA, TT} /5.0$, followed by $\epsilon^{\rm IA, TT} \rightarrow \epsilon^{\rm IA, TT}/7.0$.

%------------------------------
\subsection{HOD-NLA model}
\label{subsec:HOD}

\input{figures/shear_corr_funcs/xi_hod_nla}

\input{figures/shear_corr_funcs/xi_hod_tt}

The NLA and extended-NLA models respectively assume a null and a linear bias between the source galaxies and the underlying matter field, both of which are approximations to the galaxy dark matter connection.
In fact, galaxies populate dark matter over-densities in a more complex manner \citep{GalaxyPopulationPapers}, which is better described by the Halo Occupation Distribution approach (HOD hereafter).
The largest difference here is that these HOD galaxies are non-linear biased tracers of the matter distribution, which means that a larger number can populate regions of large over-densities, where the tidal fields are generally stronger. 
We therefore expect the impact of IA to be stronger in this case.

We investigate this effect with HOD galaxies extracted from the SkySim5000 simulation \citep{SkySim5000_HOD} \footnote{We used version: {\tt skysim5000\_v1.2}, as detailed in XYZ..}, which are obtained by [...] (mention the TXPipe paper).
Specifically, we select objects over a patch of 732.20 deg$^2$ given by  0$<$RA$<$ 20 and -36.61$<$DEC$<$0, we apply a magnitude cut of mag$_r<$24.8, and further downsample randomly to retain 20\% of the galaxies, 
therefore matching the global galaxy redshift distribution and number density found in the rest of this paper (Eq. \ref{eq:nz}).
The sample is further split into five tomographic bins by using Charlie's magical code (Charlie), resulting in the $n(z)$ presented in Fig. \ref{fig:NzHOD}. 
Compared to the analytical SRD-Y1 presented in Fig. \ref{fig:Nz}, the mean redshifts of the five tomographic bins are shifted by [-0.027, -0.018,  0.003, -0.010,  0.038], respectively.

[{\it To put somewhere?: Since source clustering is a higher order effect in cosmic shear, the measured $\xi_{\pm}$ from these three mocks (random, $b_{\rm TA}=1.0$ and $b_{\rm TA}=2.0$) show negligible differences ({\it to quantify and cite}). 
In contrast, the impact on the IA contamination is significant and can double the secondary signal in certain circumstances. 
The main explanation for this is that while clustered sources are uncorrelated with the foreground matter responsible for the lensing signal, the high-density regions they live are also subject to the largest tidal fields. 
}]

\subsection{Validation with full inference}
 \label{sec:inference}
\begin{figure}
%\includegraphics[width=\columnwidth]{graphs/triangle_nla_5D.pdf}
\includegraphics[width=\columnwidth]{graphs/triangle_bta_sweep.pdf}
\caption{Cosmological inference from the NLA-infused simulations, analysed the extended-NLA model. 
The cross-hairs represent the truth as fixed in the {\it Outer Rim} $N$-body simulations \citep{OuterRim}.}
\label{fig:triangle_nla}
\end{figure}
priors, {\sc CosmoSIS}, Takahashi $P(k)$, etc.

\input{figures/corner_plots/corner_NLA_all}

\begin{figure}
%\includegraphics[width=\columnwidth]{graphs/triangle_nla_5D.pdf}
\includegraphics[width=\columnwidth]{graphs/TT_allsweep.pdf}
\caption{Cosmological inference from the TT-infused simulations, with all 3 IA params varied. The cross-hairs represent the truth as fixed in the {\it Outer Rim} $N$-body simulations \citep{OuterRim}.}
\label{fig:triangle_tt_all_ia_params}
\end{figure}

\input{figures/corner_plots/corner_TT}