### 2024-03-27 sections/6_higher_order_wl_stats.tex

\section{Higher-order weak lensing statistics}
\label{sec:HOWLS}

... use the mocks from Table \ref{table:IAmodels} to compute derivatives. Show.