### 2024-03-27 tables/ia_models.tex

% Requires the booktabs if the memoir class is not being used
\begin{table}
   \centering
   %\topcaption{Table captions are better up top} % requires the topcapt package
    \caption{IA models infused in this work and the respective fiducial values of the model parameters.}
   \begin{tabular}{@{} lcr @{}} % Column formatting, @{} suppresses leading/trailing space
      \hline
      \hline
       %\multicolumn{2}{c}{Item} \\
      model   		& ($A_{\rm IA}, \: b_{\rm TA}, \: A_2)$ \\
      \hline
      NLA     		& {($\pm$1, 0, 0) }&  \\
      \multirow{2}{*}{extended-NLA }  	&   {($\pm$1, 1, 0)}   \\
      							&  {(1, 2, 0)}   \\
      TT 			&  {(0, 0, $\pm$1)} &  \\
      \multirow{2}{*}{extended-TT} 	&  {(0, 1 ,$\pm$1)}   \\
      				&  {(0, 2, 1)}   \\
      TATT- fiducial	&  { (1, 1, 1)}  \\
      %TATT-DESY3	&   (-0.7,1,1.9) & \\
      \hline
      \hline

   \end{tabular}
   \label{table:IAmodels}
\end{table}