### 2024-06-21 tables/ia_models.tex
Updates from Overleaf

% Requires the booktabs if the memoir class is not being used
\begin{table}
   \centering
   %\topcaption{Table captions are better up top} % requires the topcapt package
    \caption{IA models infused in this work and the respective fiducial values of the model parameters.}
   \begin{tabular}{@{} lcr @{}} % Column formatting, @{} suppresses leading/trailing space
      \hline
      \hline
      %\multicolumn{2}{c}{Item} \\
      model   		& ($A_{\rm IA}, \: b_{\rm TA}, \: A_2)$ \\
      \hline
      NLA     		& ($\pm$1, 0, 0) &  \\
      \hline
      \multirow{2}{*}{extended-NLA }  	&   ($\pm$1, 1, 0)   \\
                                        &  (1, 2, 0)   \\
      \hline
      TT 			&  {(0, 0, $\pm$1)} &  \\
      \hline
      \multirow{2}{*}{extended-TT} 	    &  (0, 1 ,$\pm$1)   \\
                                        &  (0, 2, 1)   \\
      \hline
      TATT- fiducial	&  (1, 1, 1)  \\
      \hline
      \hline
   \end{tabular}
   \label{table:IAmodels}
\end{table}