### 2024-06-21 tables/priors.tex

\begin{table}
   \centering
   \caption{Priors used in the likelihood sampling. {\it (verify these numbers with Niko)}}
   \tabcolsep=0.11cm
      \begin{tabular}{@{} cccccccc @{}} % Column formatting, @{} suppresses leading/trailing space
      \hline
      \hline
       Parameter       &  range & prior \\
       \hline
       Cosmology\\ 
       $\Omega_{\rm m}$ &[0.1, 0.55] & Flat\\
       $S_8$ & [0.6, 0.9] & Flat\\
       \hline
       IA\\
       $A_{\rm IA}$ & [-5, 5]&  Flat\\                    
        $b_{\rm TA}$ & [0, 2] & Flat\\
        $C_2$ & [-5, 5] & Flat\\
     \hline
    \hline 
    \end{tabular}
    \label{table:priors}
\end{table}